---
title: "Analysis of Optimising creativity for literary texts"
author: "Ana Guerberof Arenas"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(lme4)
library(ggplot2)
library(car)
library(ARTool)
library(glmmTMB)
library(DHARMa)
library(car)
library(gridExtra)
library(lme4)
```

# Load and Explore Data

```{r}
# Load the dataset
data <- read.csv("UCP_Annotations_All.csv")

# Display the structure of the dataset
str(data)
data$Modality <- as.factor(data$Modality)
data$UCPs <- as.factor(data$UCPs)
data$Language <- as.factor(data$Language)
levels(data$Modality)
levels(data$UCPs)
levels(data$Language)
str(data)

# Preview the data
dplyr::glimpse(data)

# Eliminate first hidden characters from UCP_ID
colnames(data)[colnames(data) == "X...UCP_ID"] <- "UCP_ID"

# Drop 3c for CA and replace 3c by 3d
# rationale. We tried 2 NMT systems (3c and 3d) and 3d is better. For the other languages we had only one (3c)
data_fil <- data %>% filter_at(vars(Language, Modality), any_vars(. != "CA" | . != "3c"))
data_fil <- data[!(data$Language == "CA" & data$Modality == "3c"), ]
data_fil$Modality[data_fil$Modality == "3d"] <- "3c"
data <- data_fil

```

# Data Preparation

```{r}
# Calculate the number of CSs (sum of CSM, CSA, CSC) per Modality
data <- data %>% 
  mutate(CS_Total = if_else(UCPs %in% c("CSM", "CSA", "CSC"), 1, 0))

# Create new column new_UCPs in case we want to compare with R, O, E.

data <- data %>% 
  mutate(new_UCPs = case_when(
    UCPs %in% c("CSM", "CSA", "CSC") ~ "CS",
    UCPs == "R" ~ "R",
    UCPs == "O" ~ "O",
    UCPs == "E" ~ "E",
    TRUE ~ NA_character_
  ))

# Summarize CSs, Errors, and Points per Modality
data_summary <- data %>% 
  group_by(Modality) %>% 
  summarize(
    Total_CS = sum(CS_Total),
    Total_Errors = sum(Errors),
    Total_Points = sum(Points),
    Total_Index = sum(Creativity_Index)
  )

data_summary

# Create subset of data without HT, to have only ChatGPT, DeepL and GT

data_filtered <- data %>%
  filter(!Modality %in% c("HT"))

# View the first few rows of the filtered dataset
head(data_filtered)

# Create subset of data with HT, to compare with best performing models 3b (ES, NL, Zh) and 1b (CA)

data_subset <- data %>%
  filter(
    (Modality %in% c("HT", "3b") & Language %in% c("ES", "NL")) | 
    (Modality %in% c("HT", "1b") & Language == "CA"))

# View the first few rows of the subset
head(data_subset)


```

# Exploratory Data Analysis

## Total CSs, Errors, and Points per Modality

```{r}
# Bar plots for CSs, Errors, and Points using the data_summary
cs_plot <- ggplot(data_summary, aes(x = Modality, y = Total_CS)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Total CSs per Modality", x = "Modality", y = "Total CSs")

errors_plot <- ggplot(data_summary, aes(x = Modality, y = Total_Errors)) +
  geom_bar(stat = "identity", fill = "tomato") +
  labs(title = "Total Errors per Modality", x = "Modality", y = "Total Errors")

points_plot <- ggplot(data_summary, aes(x = Modality, y = Total_Points)) +
  geom_bar(stat = "identity", fill = "forestgreen") +
  labs(title = "Total Points per Modality", x = "Modality", y = "Total Points")

grid.arrange(cs_plot, errors_plot, points_plot, ncol = 3)


```

## Interaction Graphics (without HT): Creative Shifts

```{r}
# Summarize CS_Total by Modality and Language
data_sum_CS <- data_filtered %>%
  group_by(Modality, Language) %>%
  summarise(Total_CS = sum(CS_Total, na.rm = TRUE))  # Sum CS_Total for each group

pdf(file="./plot_css_gpt.pdf", width = 6, height = 4)
ggplot(data_sum_CS, aes(x = Modality, y = Total_CS, color = Language, group = Language)) +
  geom_point(position = position_dodge(width = 0.3), alpha = 0.7, size = 3) +  # Ensure alignment with dodge
  geom_line(position = position_dodge(width = 0.3), alpha = 0.7, size = 1, linetype = "dotted") +  # Dodge ensures lines match points
  scale_color_brewer(palette = "Set2") +  # Use a colorblind-friendly palette
  labs(x = "Modality", 
       y = "Total CSs",
       color = "Language") +  # Change 'fill' to 'color' since we're using color mapping
  theme_minimal(base_size = 10) +  # Minimal theme with larger text
  theme(
    legend.position = "top",  # Move legend to top
    legend.title = element_text(face = "bold"),  # Bold legend title
    axis.text.x = element_text(angle = 45, hjust = 1, face = "bold"),  # Rotate and bold x-axis labels
    plot.title = element_text(face = "bold", hjust = 0.5, size = 16)  # Center and bold the title
  )
dev.off()
```

## Interaction Graphics (without HT): Error points

```{r}

# Interaction of Points and Modality

# Summarize Points by Modality and Language
data_sum_points <- data_filtered %>%
  group_by(Modality, Language) %>%
  summarise(Total_Points = sum(Points, na.rm = TRUE))  # Sum Points for each group

# Calculate the overall mean of Total Points
#mean_points <- mean(data_sum_points$Total_Points, na.rm = TRUE)

pdf(file="./plot_errors_gpt.pdf", width = 6, height = 4)
ggplot(data_sum_points, aes(x = Modality, y = Total_Points, color = Language, group = Language)) +
  geom_point(position = position_dodge(width = 0.3), alpha = 0.7, size = 3) +  # Ensure alignment with dodge
  geom_line(position = position_dodge(width = 0.3), alpha = 0.7, size = 1, linetype = "dotted") +  # Dodge ensures lines match points
  scale_color_brewer(palette = "Set2") +  # Use a colorblind-friendly palette
  labs(x = "Modality", 
       y = "Total Error points",
       color = "Language") +  # Change 'fill' to 'color' since we're using color mapping
  theme_minimal(base_size = 10) +  # Minimal theme with larger text
  theme(
    legend.position = "top",  # Move legend to top
    legend.title = element_text(face = "bold"),  # Bold legend title
    axis.text.x = element_text(angle = 45, hjust = 1, face = "bold"),  # Rotate and bold x-axis labels
    plot.title = element_text(face = "bold", hjust = 0.5, size = 16)  # Center and bold the title
  )
dev.off()
```


## Interaction Graphics subset of data: best models with HT

```{r}
# Interaction of CS_Total and Modality in subset

# Summarize the data: the sum of CS_Total for each Modality and Language
data_subset_sum_CS <- data_subset %>%
  group_by(Modality, Language) %>%
  summarize(CS_Total = sum(CS_Total, na.rm = TRUE))

data_subset_points <- data_subset %>%
  group_by(Modality, Language) %>%
  summarize(Total_Points = sum(Points, na.rm = TRUE))

# Create the plot with the summarized data
pdf(file="./plot_css_ht.pdf", width = 6, height = 4)
ggplot(data_subset_sum_CS, aes(x = Modality, y = CS_Total, fill = Language)) +
  geom_col(position = "dodge") +  # Side-by-side bars for each Language
  scale_fill_brewer(palette = "Set2") +  # Use the same colorblind-friendly palette
  labs(x = "Modality", 
       y = "Total CSs", 
       fill = "Language") +
  theme_minimal(base_size = 10) +  # Keep the same text size
  theme(
    legend.position = "top",  # Move legend to the top for consistency
    legend.title = element_text(face = "bold"),  # Bold legend title
    axis.text.x = element_text(angle = 45, hjust = 1, face = "bold"),  # Rotate and bold x-axis labels
    )
dev.off()

# Interaction of Points and Modality and Language
pdf(file="./plot_errors_ht.pdf", width = 6, height = 4)
ggplot(data_subset_points, aes(x = Modality, y = Total_Points, fill = Language)) +
  geom_col(position = "dodge") +  # Side-by-side bars for each Language
  scale_fill_brewer(palette = "Set2") +  # Use the same colorblind-friendly palette
  labs(x = "Modality", 
       y = "Total Error Points", 
       fill = "Language") +
  theme_minimal(base_size = 10) +  # Keep the same text size
  theme(
    legend.position = "top",  # Move legend to the top for consistency
    legend.title = element_text(face = "bold"),  # Bold legend title
    axis.text.x = element_text(angle = 45, hjust = 1, face = "bold"),  # Rotate and bold x-axis labels
    )
dev.off()
```


# Modeling

## Interaction of CSs per Modality (with Repeated Measures)

```{r}
# Mixed model for CSs
cs_model <- lmer(CS_Total ~ Modality + (1|UCP_ID) + (1|Language) , data = data_filtered)
summary(cs_model)

# Check assumptions
plot(resid(cs_model), fitted(cs_model), main = "Residuals vs Fitted: CS Model")
hist(resid(cs_model), main = "Residuals Histogram: CS Model", xlab = "Residuals")
qqnorm(resid(cs_model))
qqline(resid(cs_model))

# Assumptions do not check

# Transform data to avoid 0s and 1s (Beta regression requires (0,1) range)
data_filtered$CS_Total_adj <- (data_filtered$CS_Total * (nrow(data_filtered) - 1) + 0.5) / nrow(data_filtered)

# Fit the Beta GLMM
model <- glmmTMB(CS_Total_adj ~ Modality * Language + (1 | UCP_ID), 
                 family = beta_family(), 
                 data = data_filtered)

# Check model summary
summary(model)

# Simulate residuals for GLMM
res_glmm <- simulateResiduals(model)

# Residual diagnostics plot
plot(res_glmm)

testDispersion(res_glmm)

# Initial assumptions do not check.

# Mixed model for Points
points_model <- lmer(Points ~ Modality * Language + (1|UCP_ID), data = data_filtered)
summary(points_model)

# Check assumptions

plot(residuals(points_model) ~ fitted(points_model))
hist(residuals(points_model))  # Histogram
qqnorm(residuals(points_model))
shapiro.test(residuals(points_model))
acf(residuals(points_model))
ranef(points_model)
vif(points_model)

# Assumptions do not check

```

# Non-parametric tests

## Interaction of CSs per Modality (with Repeated Measures)

```{r}

# Non parametric Tests. 
# Since the assumptions are not met in the model, We will use ART ANOVA. The Aligned Rank Transform (ART) ANOVA is a non-parametric alternative to standard ANOVA that allows for testing main effects and interactions in factorial designs.

# We use only the data_filtered here

model_art_CS <- art(CS_Total ~ Modality * Language + (1 | UCP_ID), data = data_filtered)
anova(model_art_CS)


# Pairwise comparisons for Modality
art.con(model_art_CS, "Modality", adjust = "bonferroni") 

# Pairwise comparisons for Language
art.con(model_art_CS, "Language", adjust = "bonferroni") 

# Pairwise comparisons for Interaction (Modality * Language)
art.con(model_art_CS, "Modality:Language", adjust = "bonferroni")


```


## Interaction of Points per Modality (with Repeated Measures)

```{r}

# Run also ART ANOVA (non-parametric) for the points
art_model_points <- art(Points ~ Modality * Language + (1|UCP_ID), data = data_filtered)

# ANOVA
anova(art_model_points)


# Pairwise comparisons for Modality
art.con(art_model_points, "Modality", adjust = "bonferroni") 
# Pairwise comparisons for Language
art.con(art_model_points, "Language", adjust = "bonferroni") 

# Pairwise comparisons for Interaction (Modality * Language)
art.con(art_model_points, "Modality:Language", adjust = "bonferroni")

# Run ART ANOVA (non-parametric) only Modality to double check

art_model_modality <- art(Points ~ Modality + (1|UCP_ID), data = data_filtered)
anova(art_model_modality)

# Pairwise comparisons for Modality
art.con(art_model_modality, "Modality", adjust = "bonferroni")


```


## HT vs MT and Creative Shifts

```{r}
# Create a dataset to compare the best options with HT from data_subset
data_subset <- data_subset %>% 
  mutate(new_Modality = case_when(
    Modality == "HT" ~ "HT",
    Modality == "3b" ~ "MT",
    Modality == "1b" ~ "MT",
    TRUE ~ NA_character_
  ))

data_subset$new_Modality <- as.factor(data_subset$new_Modality)

# Run art anova for CSs

art_model_HT_CS <- art(CS_Total ~ new_Modality * Language + (1 | UCP_ID), data = data_subset)
anova(art_model_HT_CS)

# Pairwise comparisons for Modality
art.con(art_model_HT_CS, "new_Modality", adjust = "bonferroni") 

# Pairwise comparisons for Language
art.con(art_model_HT_CS, "Language", adjust = "bonferroni") 

# Pairwise comparisons for Interaction (Modality * Language)
art.con(art_model_HT_CS, "new_Modality:Language", adjust = "bonferroni")
```


## HT vs MT and Error points

```{r}
## Interaction of CS per Modality, HT vs MT

# Run ART ANOVA (non-parametric)
art_model_HT <- art(Points ~ new_Modality * Language + (1|UCP_ID), data = data_subset)

# ANOVA for ART model
anova(art_model_HT)

# Pairwise comparisons for Modality
art.con(art_model_HT, "new_Modality", adjust = "bonferroni") #No significance

# Pairwise comparisons for Language
art.con(art_model_HT, "Language", adjust = "bonferroni") #Significance CA with other languages

# Pairwise comparisons for Interaction (Modality * Language)
art.con(art_model_HT, "new_Modality:Language", adjust = "bonferroni") #No significance
```





